\begin{defn}[Category]
A \emph{category} $\mathscr{A}$ consists of
\begin{itemize}
  \item{
    a collection of objects $\mathrm{Ob}(\mathscr{A})$
  }
  \item{
    for any two \emph{objects} $A, B \in \mathrm{Ob}(\mathscr{A})$
    a set (the hom-set) of \emph{morphisms} or \emph{arrows}
    $\mathrm{Mor}(A, B)$ from $A$ to $B$
  }
  \item{
    for any three objects $A$, $B$, $C$ a \emph{composition law}
    $$
    \circ : \mathrm{Mor}(A, B) \times \mathrm{Mor}(B, C) \to
    \mathrm{Mor}(A, C)
    $$
    given by $(f, g) \mapsto g \circ f$ such that
    \begin{enumerate}
      \item{
        $\mathrm{Mor}(A, B) \cap \mathrm{Mor}(A^\prime, B^\prime) =
        \varnothing$ unless $A = A^\prime$, $B = B^\prime$
      }
      \item{
        for each $A \in \mathrm{Ob}(\mathscr{A})$ a morphism
        $\mathrm{id}_A \in \mathrm{Mor}(A, A)$ such that
        $\forall f \in \mathrm{Mor}(A, B)$, $f \circ \mathrm{id}_A =
        f$
        and $\forall g \in \mathrm{Mor}(B, A)$
        $\mathrm{id}_A \circ g = g$
      }
      \item{
        for $A, B, C, D \in \mathrm{Ob}(\mathscr{A})$ the
        associativity law $(h \circ g) \circ f = h \circ (g \circ f)$.
      }
    \end{enumerate}
  }
\end{itemize}
\end{defn}

\begin{xmpl}
\begin{enumerate}
  \item{
    The category $\mathbf{Set}$ with objects sets and morphisms functions.
  }
  \item{
    The category $\mathbf{Grp}$ with objects groups and morphisms
    group homomorphisms. This has a subcategory $\mathbf{Ab}$ of
    abelian groups with group homomorphisms.
  }
  \item{
    The category $\mathbf{Top}$ with objects topological spaces and
    morphisms continuous maps.
  }
  \item{
    A directed graph with objects vertices and hom-sets
    $\mathrm{Mor}(v_1, v_2)$ the set of all paths from $v_1$ to $v_2$.
  }
  \item{
    Let $D \subset \mathbb{R}^2$ (more generally some topological
    space). Objects are taken to be the points of $D$, and for points
    $p, q$ the morphisms $\mathrm{Mor}(p, q)$ are all paths in $D$ that start
    from $p$ and end at $q$, i.e. continuous maps $\gamma : [0,1] \to
    D$ such that $\gamma(0) = p$, $\gamma(1) = q$, up to
    reparametrization. That is, $f_1$ and $f_2$ give the same path if
    $\exists g : [0,1] \to [0,1]$ bijective, bicontinuous such that
    $f_2 = f_1 \circ g$. Here composition of $f \in \mathrm{Mor}(p,q)$
    and $g \in \mathrm{Mor}(q,r)$ is the equivalence class (up to
    reparametrization) of
    $$
    h(x) =
    \left\{\begin{array}{c c}
      f(2x),   & x \in [0, 1/2] \\
      g(2x-1), & x \in [1/2, 1]
    \end{array}\right.
    $$
  }
\end{enumerate}
\end{xmpl}

\begin{defn}[Functors]
Let $\mathscr{A}$, $\mathscr{B}$ be categories. A functor
$F : \mathscr{A} \to \mathscr{B}$ consists of
\begin{itemize}
  \item{
    a rule that to each $A \in \mathrm{Ob}(\mathscr{A})$ associates
    $F(A) \in \mathrm{Ob}(\mathscr{B})$
  }
  \item{
    for each $A$, $A^\prime$ objects in $\mathscr{A}$
    a map $\mathrm{Mor}(A, A^\prime) \to \mathrm{Mor}(F(A),
    F(A^\prime))$ denoted $f \mapsto F(f)$ such that
    \begin{itemize}
      \item{
        for all $A \in \mathrm{Ob}(\mathscr{A})$,
        $F(\mathrm{id}_A) = \mathrm{id}_{F(A)}$
      }
      \item{
        for all $A, B, C \in \mathrm{Ob}(\mathscr{A})$
        and all $f : A \to B$, $g : B \to C$ we have
        $F(g \circ f) = F(g) \circ F(f)$.
      }
    \end{itemize}
  }
\end{itemize}
\end{defn}

\begin{xmpl}
\begin{enumerate}
  \item{
    Let $\mathscr{A} = \mathbf{Grp}$, $\mathscr{B} = \mathbf{Set}$.
    The \emph{forgetful functor} $F$ takes a group $G$
    to the underlying set $F(G)$ and the group homomorphism
    $f : G \to H$ to $F(f) : F(G) \to F(H)$ the set underlying set map.
  }
  \item{
    There is a functor $F : \mathbf{Grp} \to \mathbf{Ab}$ that takes
    $G \mapsto G^{ab} = G / [G, G]$, the maximal abelian quotient.
    Here $[G, G]$ is the subgroup of $G$ generated by
    $[a, b] = a b a^{-1} b^{-1}$. This is the maximal quotient of $G$
    that is abelian, i.e. if $f : G \to H$ with $H$ abelian, then
    there exists a unique $\bar{f} : G / [G, G] \to H$ such that
    $f$ factors through this map, i.e. $G \to G / [G,G] \to H$.

    To confirm that this is a functor we need to check that
    for all group homomorphisms $f : G \to H$ there is an
    induced group homomorphism $F(f) : G^{ab} \to H^{ab}$. This is
    because $f([G, G]) \subset [H, H]$ since
    $f(a b a^{-1} b^{-1}) = f(a)f(b)f(a)^{-1}f(b)^{-1}$.
  }
  \item{
    There is a functor $F : \mathbf{Set} \to \mathbf{Grp}$ that takes
    $S \mapsto F(S)$, the free group generated by $S$. If we have $f :
    S \to T$ a set map induces group homomorphism
    $F(f) : F(S) \to F(T)$, e.g.
    $$
    F(f)(a^2b^3c^4) = f(a)^2f(b)^3f(c)^4.
    $$
    We can also do this by employing the universal property of the
    free group:
   $$
   \begindc{\commdiag}[500]
   \obj(0,0)[S]{$S$}
   \obj(1,0)[FS]{$F(S)$}
   \obj(0,1)[T]{$T$}
   \obj(1,1)[FT]{$F(T)$}
   \mor{S}{T}{$f$}
   \mor{S}{FS}{$i$}
   \mor{T}{FT}{$i$}
   \mor{S}{FT}{$\bar{f}$}[\atright, \dashArrow]
   \mor{FS}{FT}{$F(f)$}[\atright, \dashArrow]
   \enddc
   $$
  }
\end{enumerate}
\end{xmpl}

\begin{defn}[Isomorphism]
Let $\mathscr{A}$ be a category, $A$, $B$ objects, and
$f \in \mathrm{Mor}(A, B)$. $f$ is an isomorphism in
$\mathscr{A}$ when there exists a $g \in \mathrm{Mor}(B, A)$ such that
$$
f \circ g = \mathrm{id}_B, \quad
g \circ f = \mathrm{id}_A.
$$
\end{defn}
