\section{Groups}

\subsection{Definition}

\begin{defn}[Group]
  \label{def:group}
  A \emph{group} $(G, \cdot)$ is a monoid such that all elements have an inverse.
  That is to say
  \begin{enumerate}
    \item{$G$ is a set,
         }
    \item{$\cdot : G \times G \to G$ is an associative binary operation on $G$,
         }
    \item{There exists a unique $1 \in G$ (sometimes denoted $e$) such that
          $1 \cdot g = g \cdot 1 = g, \forall g \in G$,
         }
    \item{For every $g \in G$, there exists a unique $g^{-1} \in G$ such that
          $g \cdot g^{-1} = g^{-1} \cdot g = e$.
         }
  \end{enumerate}
  The \emph{order} of the group is the cardinality of $G$. The order
  of an element $a \in G$ is the smallest positive integer $m$ such
  that $a^m = e$.
\end{defn}

\begin{xmpl}
  Some examples of groups:
  \begin{enumerate}
    \item{The integers $\mathbb{Z}$ form a group with $\cdot = +$.}
    \item{The rationals $\mathbb{Q}$ form a group with $\cdot = +$.}
    \item{The reals except zero $\mathbb{R}^\ast = \mathbb{R} - {0}$
          form a group with respect to multiplication.}
    \item{Given a set $X$,
          $$
          \mathrm{Perm}(X) = \{f : X \to X | f \text{ is bijective}\}
          $$
          is a group with respect to composition ($\cdot = \circ$).
          }
     \item{When $X$ is a finite set $X = \{1, \dots, n\}, n \in \mathbb{N}$,
           then $\mathrm{Perm}(X) = S_n$, the symmetric group with $n$ letters.
           The order of $S_n$ is $n!$.
          }
     \item{For $n \geq 1$, the set of $n \times n$ matrices $A$
           with real entries such that $\det(A) \neq 0$ is a
           group with respect to multiplication of matrices, denoted
           $GL_n(\mathbb{R})$ (the \emph{general linear group}).
           It is closed under this operation since
           $\det(A \cdot B) = \det(A) \det(B) \neq 0$. This product is
           associative since linear maps are associative, and thus so is
           multiplication of their corresponding matrices. It possesses
           an identity element $e = I_n$ and inverses
           $A^{-1} = \frac{1}{\det A} A^{\mathrm{adj}}$, where
           the $(i, j)$ element of $A^{\mathrm{adj}}$ is $(-1)^{i + j}$ times
           the minor determinant of the $(n-1) \times (n-1)$ matrix obtained
           by erasing the $j$-th row and the $i$-th column.
           Recall that in the $2 \times 2$ case
           $$
           \left[\begin{array}{c c}
             a & b \\ c & d
           \end{array}\right]^{-1}
         = \frac{1}{ad - bc}
           \left[\begin{array}{r r}
             d & -b \\ -c & a
           \end{array}\right].
           $$
           }
  \end{enumerate}
\end{xmpl}

\begin{defn}[Abelian Group]
  \label{def:abelian-group}
  A group $(G, \cdot)$ is \emph{abelian} or \emph{commutative} when
  $\forall x, y \in G$, $x \cdot y = y \cdot x$.
\end{defn}

\begin{Problem}
  For $n > 1$, show $\mathrm{GL}_n(\mathbb{R})$ is not abelian.
  For $n > 2$, show $S_n$ is not abelian.
  Hint: It is enough to show that $S_3$ is not abelian because
  $S_n$ is a subgroup of $S_{n+1}$ for all $n$.
\end{Problem}

\begin{Answer}
Consider the $n \times n$ matrices $A_{ij} = \delta_{i1}$ and
$B_{ij} = \delta_{1j}$. Then
\begin{align*}
(A B)_(ij)
&= \sum_{k=1}^n A_{ik} B_{kj} = \sum_{k=1}^n \delta_{i1}\delta{1j} \\
&= \left\{\begin{array}{c c}
     n, & \quad i = j = 1 \\
     0, & \quad \text{otherwise}
\end{array}\right.
\end{align*}
while
\begin{align*}
(B A)_(ij)
&= \sum_{k=1}^n B_{ik} A_{kj} = \sum_{k=1}^n \delta_{1k}\delta{k1} \\
&= 1.
\end{align*}

In $S_3$ we observe that
$$
(1 2 3) (1 2) = (1 3)
$$
while
$$
(1 2) (1 2 3) = (2 3).
$$
\end{Answer}

\begin{defn}[Subgroup]
  \label{def:subgroup}
  A subset $H \subset G$ is called a \emph{subgroup} of $G$ when
  \begin{enumerate}
    \item{$e \in H$,}
    \item{$x, y \in H \implies x \cdot y \in H$,}
    \item{$x \in H \implies x^{-1} \in H.$}
  \end{enumerate}
  We write $H < G$.
\end{defn}

\begin{xmpl}
  \item{$\mathbb{Z} < \mathbb{Q}$ with $+$.}
  \item{$\{A \in \mathrm{GL}_n(\mathbb{R}) | \det A = 1\} < \mathrm{GL}_n(\mathbb{R})$
        because $\det (A \cdot A^{-1}) = \det I = 1$,
        so $\det(A) = 1 \implies \det(A^{-1}) = 1$. This is the
        \emph{special linear group} $\mathrm{SL}_n(\mathbb{R})$.
       }
\end{xmpl}

\begin{defn}[Direct Product Group]
  Given groups $G_1, G_2$,
  $$
  G_1 \times G_2 = \{(g_1, g_2) | g_1 \in G_1, g_2 \in G_2 \}
  $$
  is a group, with product given by
  $$
  (g_1, g_2) \cdot (g_1^\prime, g_2^\prime)
  = (g_1 \cdot_1 g_1^\prime, g_2 \cdot_2 g_2^\prime),
  $$
  called the \emph{direct product group}.
\end{defn}

% 9/5
\subsection{Generalities}
Let $G$ be a group, and let $S \subset G$. Consider
$$
\langle S \rangle = \bigcap_{\substack{H < G \\ S \subset H}} H,
$$
the intersection of all subgroups that contain $S$.

\begin{prop}
$\langle S \rangle$ is a subgroup of $G$, and is the smallest
subgroup of $G$ that contains $S$. Every element of $\langle S \rangle$
can be written as a product $x_1 \cdots x_n$ with either $x_i \in S$ or
$x_i^{-1} S$, where the empty product is taken to be 1. Then
$\langle S \rangle \equiv$ the subgroup of $G$ generated by $S$.
\end{prop}

\begin{proof}
First, every intersection of any family of subgroups of a group is a
subgroup (Exercise). Observe $S \subset \langle S \rangle$, and also if
$H^\prime < G$ such that $S \subset H^\prime$ then
$\langle S \rangle \subset H^\prime$, since
$\langle S \rangle = \bigcap_{\substack{H < G \\ S \subset H}} H \subset H^\prime$.

Next, to show
$\langle S \rangle = \{ x_1 \cdots x_n | x_i \in S \lor x_i^{-1} \in S \},$
observe that $\{ x_1 \cdots x_n | x_i \in S \lor x_i^{-1} \in S \} < G$ and
contains $S$, and in fact each subgroup of $G$ that contains $S$ has to
contain this set, and thus so does their intersection.
\end{proof}

\begin{Problem}
Every intersection of a family of subgroups of a group is a subgroup.
\end{Problem}
\begin{Answer}
Let $\{ H_i \}_{i \in I}$ be a family of subgroups and let
$x, y \in H = \bigcap_{i \in I} H_i$. Then for each $i \in I$,
$x, y \in H_i$, so $xy \in H_i$ since $H_i$ is a subgroup. But
$xy \in H_i$ for all $i \in I$ means that $xy \in H$. Similarly $x \in
H$ gives $x^{-1} \in H$, so $H$ is a subgroup.
\end{Answer}

\begin{defn}[Generating Set, Cyclic Group]
  We define what is meant by a generator.
  \begin{enumerate}
    \item{If for some $S \subset G$, $\langle S \rangle = G$, we say that
          $S$ \emph{generates} $G$.
         }
    \item{If $G$ is generated by a single element $a \in G$, we say $G$ is
          \emph{cyclic} with generator $A$. Then
          $$
          G = \langle a \rangle = \{ \dots, a^{-2}, a^{-1}, 1, a, a^2, \dots \}.
          $$
         }
  \end{enumerate}
\end{defn}

\subsection{Homomorphisms}

\begin{defn}[Group Homomorphism]
For $G, G^\prime$ groups, a map $f : G \to G^\prime$ is a \emph{homomorphism}
when $\forall x, y \in G$, $f(x \cdot y) = f(x) \cdot f(y)$. Then
$$
f(e) = f(e \cdot e) = f(e) \cdot f(e)
\implies e^\prime = f(e)^{-1} \cdot f(e) \cdot f(e) = f(e)
$$
and
$$
e^\prime = f(e) = f(x \cdot x^{-1}) = f(x) \cdot f(x^{-1})
\implies f(x^{-1}) = f(x)^{-1},
$$
so a group homomorphism preserves identities and inverses.
\end{defn}

\begin{prop}
Given homomorphisms $f : G \to G^\prime$
and $g : G^\prime \to G^{\prime\prime}$,
$g \circ f$ is a homomorphism.
\end{prop}

\begin{proof}
We have
$$
(g \circ f)(xy) = g(f(x)f(y))
                = g(f(x))g(f(y))
                = (g \circ f)(x) (g \circ f)(y)
$$
\end{proof}

\begin{defn}[Group Isomorphism]
A bijective group homomorphism is called a \emph{group isomorphism}.
Then the inverse map $f^{-1} : G^\prime \to G$ is also a group
isomorphism since each $x^\prime \in G^\prime$ is given by $x^\prime = f(x)$
for some $x \in G$, so
$$
f^{-1}(x^\prime y^\prime) = f^{-1}(f(x)f(y))
                      = f^{-1}(f(xy))
                      = x y
                      = f^{-1}(x^\prime) f^{-1}(y^\prime).
$$
We say that groups $G, G^\prime$ are isomorphic if there exists an isomorphism
between them. We write $G \simeq G^\prime$.
\end{defn}
\begin{xmpl}
For $n \geq 2$, $\mathbb{Z}_n = \{ 0, 1, \dots, n-1 \}$ (the integers mod $n$)
with addition
and $G^\prime = \{ z \in \mathbb{C} | z^n = 1 \}$ with multiplication
are isomorphic, with an isomorphism $f$ given by
$$
f(a \mod n) = e^{\frac{2 \pi i a}{n}}.
$$
We verify that
\begin{align*}
f((a \mod n) + (b \mod n))
 &= f(a + b \mod n)
  = e^{\frac{2 \pi i (a + b)}{n}} \\
 &= e^{\frac{2 \pi i a}{n}} e^{\frac{2 \pi i b}{n}} \\
 &= f(a \mod n) f(b \mod n)
\end{align*}
and that since $0 \leq a \leq n-1$, the argument
$0 \leq \frac{2 pi a}{n} < 2\pi$ so
with an appropriate choice of branch cut we have
$$
f^{-1}(z) = \frac{n}{2 \pi i} \log z \mod n.
$$
\end{xmpl}

Consider the automorphisms
$\mathrm{Aut}(G) = \{f : G \to G | f \text{ is an isomorphism}\}$.
Then $\mathrm{Aut}(G)$ is a group with $\circ$. This is a subgroup
of $\mathrm{Perm}(G)$.

\begin{Problem}
What is $\mathrm{Aut}(\mathbb{Z}_n)$?
Hint: $f : \mathbb{Z}_n \to \mathbb{Z}_n$ is determined by $f(1 \mod n)$.
Consider integers prime to $n$.
\end{Problem}

\begin{Answer}
We note that since
$$
f(m) = f\left(\sum_{k=1}^m 1\right) = \sum_{k=1}^m f(1) = m \cdot f(1),
$$
any homomorphism $f : \mathbb{Z}_n \to \mathbb{Z}_n$ is determined by
its action on $1$. Furthermore, for such an endomorphism to have an
inverse $g$ which is a homomorphism, we muust have that
$$
g(f(1)) = g\left(\sum_{k=1}^{f(1)} 1\right)
        = \sum_{k=1}^{f(1)} g(1)
        = f(1) \cdot g(1),
$$
and therefore we require that $g(1)$ is the multiplicative inverse of
$f(1)$. But the units in $\mathbb{Z}_n$ are exactly the integers
coprime to $n$, since the condition that $m$ coprime to $n$ is
equivalent to the existence of an integer $y$ such that $my \equiv 1
\mod n$,
so $f(1)$ must be coprime to $n$ for any automorphism
$f : \mathbb{Z}_n \to \mathbb{Z}_n$. This fact of modular arithmetic
follows from the identity $\mathrm{gcd}(m, n) = mx + by$ for some
integers $x$ and $y$.

Sidenote: Consider the group action
$\varphi : \mathbb{Z}_+ \to \mathrm{Perm}(\mathbb{Z})$ given by
$\varphi(n) x = x \bmod n$. Note that the kernel of $\varphi(n)$
is the set of all $x \in \mathbb{Z}_+$ such that $x \bmod n = 0$,
i.e. the multiples of $n$ or the ideal $n \mathbb{Z}$. Then from the
first isomorphism theorem we know that
$$
\mathbb{Z} / n \mathbb{Z}
=      \ker \varphi(n)
\simeq \mathrm{Im} \varphi(n)
=      \{0, 1, 2, \dots, n-1\}
=      \mathbb{Z}_n.
$$
\end{Answer}

\begin{defn}
Let $f: G \to G^\prime$ be a group homomorphism. Then the \emph{kernel} of
$f$ $\ker(f) \triangleq \{ x \in G | f(x) = e^\prime \}$ and the \emph{image} of $f$
$\mathrm{Im}(f) \triangleq \{ y \in G^\prime | \exists x \in G . f(x) = y \}$.
\end{defn}

\begin{prop}
$\ker(f), \mathrm{Im}(f) < G$.
\end{prop}
\begin{proof}
\begin{align*}
  x, y \in \ker(f)
    & \implies f(x) = e^\prime, f(y) = e^\prime \\
    & \implies f(x \cdot y) = f(x) \cdot f(y)
                            = e^\prime \cdot e^\prime
                            = e^\prime          \\
    & \implies x \cdot y \in \ker(f).
\end{align*}
Similarly $x \in \ker(f) \implies x^{-1} \in ker(f)$,
and $e \in ker(f)$ since homomorphisms preserve identity.
\end{proof}

\begin{prop}
$\ker f$ is trivial iff. $f$ is injective.
\end{prop}
\begin{proof}
$$
f(x) = f(y) \iff f(x) f(y)^{-1} = e^\prime
            \iff f(xy^{-1}) = e^\prime
            \iff xy^{-1} \in \ker f.
$$
Thus if $\ker f = \{ e \}$ then $x = y$. For the converse, let $f$ be
injective, and let $x \neq e$. Then $f(x) \neq f(e) = e$, so
$f(x) \notin \ker f$. Therefore $\ker f = \{ e \}$.
\end{proof}

\begin{xmpl}
Consider the groups
$G = \mathrm{GL}_n (\mathbb{R})$,
$G^\prime = \mathbb{R}^\ast = \mathbb{R} - \{0\}$ with multiplication.
$\det$ is a group homomorphism, and
$\ker(\det) = \mathrm{SL}_n(\mathbb{R})$.
\end{xmpl}

\subsection{Cosets}

\begin{defn}[Coset]
Let $H < G$. A \emph{left coset} of $H$ in $G$ is a subset in $G$ of the form
$$
aH \triangleq \{ ah | h \in H \}
$$
for some $a \in G$. A \emph{right coset} is defined similarly.
\end{defn}

\begin{remark}

Properties of cosets:
\begin{enumerate}
  \item{When $h \in H$, we have $hH = H$. If $g \in hH$ then
        $g = h h^\prime \in H$ so $g \in H$. If $h \in H$ then
        $h = h^\prime (h^\prime)^{-1} h \in h^\prime H$ for any
        $h^\prime \in H$.
       }
  \item{Two left cosets are either disjoint or identical. Take $aH$,
        $a^\prime H$ for $a, a^\prime \in G$. Suppose
        $aH \cap a^\prime H \neq \varnothing$. Then
        $\exists x \in aH \cap a^\prime H$ so that
        $x = ah_1 = a^\prime h_2$, $h_1, h_2 \in H$ so
        $a^\prime = a h_1 h_2^{-1} \implies a^\prime \in aH$,
        so $a^\prime h \in aH$, $\forall h \in H$. Thus
        $a^\prime H \subset aH$. Symmetrically
        $aH \subset a^\prime H$.
        Thus $aH = a^\prime H$.
       }
  \item{ There is a bijection $H \to aH$ given by $h \mapsto ah$. Thus
         all cosets have the same cardinality as $H$.
       }
\end{enumerate}

\end{remark}

\begin{defn}[Subgroup Index]
The number of left cosets of a subgroup $H$ in $G$, called the \emph{index} of
$H$ in $G$, is denoted $[G:H]$. The order of a group may alternatively
be defined as the index of its trivial subgroup $[G : 1]$.
\end{defn}

\begin{corol}
Some results about subgroup indices:
\begin{enumerate}
  \item{$G$ is the disjoint union of all its left cosets.}
  \item{If $G$ is finite (meaning the underlying set is finite), then
        $|G| = [G:H] |H|$.}
  \item{If $G$ is finite and $H < G$, the order of $H$ divides the
        order of $G$ (Lagrange's theorem).}
\end{enumerate}
\end{corol}

As an important consequence of the last point, note that if $H$ and
$K$ are subgroups then $H \cap K$ is a subgroup of both $H$ and
$K$. Suppose that a group $G$ has two subgroups
$P_1$, $P_2$ of identical prime order $p$. Then Lagrange's theorem says that
$|P_1 \cap P_2|$ divides $p$, so either
$P_1 \cap P_2 = \{ 1 \}$ or $P_1 \cap P_2 = p$. That is, two
subgroups of identical prime order are either identical or disjoint
except for the identity.

\begin{proof}

\begin{enumerate}
  \item{
    Let $H < G$.
    Clearly each coset of a subgroup $H < G$ is a subset of $G$, and
    likewise so is their union Let
    $g \in G$. Then $g = g \cdot 1$, but $1 \in H$ since $H$ is a
    subgroup. Therefore $g \in gH$, so
    $G = \cup_{g \in G} gH$.
    As shown above the cosets of $H$
    are pairwise disjoint, so this means $G = \coprod_{g \in G} gH$
    where $g$ is a representative of each coset.
  }
  \item{
    Lagrange's theorem and the more general formula
    $$
    [G : K] = [G : H] [H : K]
    $$
    (which is as above when $K = \{ 1 \}$) can be proved by showing
    that for coset decompositions
    $$
    H = \coprod_i x_i K, \quad
    G = \coprod_j y_j H
    $$
    we have
    $$
    G = \coprod_{i,j} y_j x_i K,
    $$
    where $x_i \in H$, $y_j \in G$, i.e. we wish to show that
    the representatives $y_j x_i$ of cosets of $K$ in $G$ are
    disjoint.

    Suppose
    $$
    y_j x_i K = y_{j^\prime} x_{i^\prime} K.
    $$
    Then since $K \subset H$ and $x_i, x_{i^\prime} \in H$,
    $y_j x_i K = y_j H$ and
    $y_{j^\prime} x_{i^\prime} K = y_{j^\prime} H$, so
    $y_j H = y_{j^\prime} H$. But by assumption $G$ is decomposed as a
    disjoint union of cosets of $H$ with $y_j$, $y_{j^\prime}$ as
    represenatives, so these cosets can only
    be the same if $y_j$ and $y_{j^\prime}$ are the same
    representative, i.e. $y_j = y_{j^\prime}$. Then multiplying
    through our original equation on the left gives
    $x_i K = x_{i^\prime} K$, and so by the same argument this means
    $x_i = x_{i^\prime}$. Therefore we have
    $$
    y_j x_i K = y_{j^\prime} x_{i^\prime} K
    \implies
    y_j x_i = y_{j^\prime} x_{i^\prime}
    $$
    which means that $y_j x_i$ are disjoint representatives of the
    coset decomposition for $K$ in $G$, as desired.

    It follows that
    $$
    G = \coprod_{i,j} y_j x_i K
      = \coprod_j y_j \left( \coprod_i x_i K \right),
    $$
    where the inner coproduct is over $[H : K]$ elements and the
    outer coproduct is over $[G : H]$, so it follows that the total
    number of elements $[G : K]$ in this coset decomposition is
    $[G : H][H : K]$.
  }
  \item{
    We can now furthermore show that $|S_n| = n!$ by
    induction. Since $S_1 = \{ \mathrm{id} \}$ we have
    $S_1 = 1$. Let $H < S_n$ be the set of permutations that hold $n$
    fixed. This is the set of permutations of $n-1$ letters so
    $|H| = |S_{n-1}|$. Now for $i \in \{1, \dots, n\}$ let $\sigma_i$
    be any permutation such that $\sigma_i(n) = i$.

    Let $\sigma \in S_n$ such that $\sigma(n) = k$. Then we observe that
    $$
    \tau(x) =
    \left\{\begin{array}{c c}
      n,                    & \quad x =    n \\
      \sigma_k^{-1}\sigma(x) & \quad x \neq n
    \end{array}\right.
    $$
    leaves $n$ fixed and so $\tau \in H$, and furthermore that
    $$
    \sigma_k \tau(x) =
    \left\{\begin{array}{c c}
      \sigma_k(n) = k = \sigma(n),               & \quad x = n \\
      \sigma_k\sigma_k^{-1}\sigma(x) = \sigma(x), & \quad x \neq n
    \end{array}\right. = \sigma(x)
    $$
    so $\sigma = \sigma_k \tau \in \sigma_k H$. Therefore
    $S_n \subset \cup_{i=1}^n \sigma_i H$.

    Furthermore let $\sigma_i H = \sigma_j H$ for some $i, j$. Then
    for some $\tau, \tau^\prime \in H$ we have
    $\sigma_i \tau = \sigma_j \tau^\prime$. Then
    $\sigma_i \tau (n) = \sigma_i (n)$ and
    $\sigma_j \tau^\prime (n) = \sigma_j (n)$
    since elements of $H$ leave $n$ fixed, so
    $\sigma_i(n) = \sigma_j(n)$. But $\sigma_i$ and $\sigma_j$ fix all
    elements other than $n$ and so this means $\sigma_i =
    \sigma_j$. Therefore $\sigma_i H$ and $\sigma_j H$ are either
    disjoint or equal.

    We conclude that
    $$
    S_n = \coprod_{i=1}^n \sigma_i H
    $$
    and so
    $$
    |S_n| = n |H| = n |S_{n-1}|.
    $$
    Therefore $|S_n| = n!$.

    Note that this argument is equivalent to counting by
    group actions, since we can instead regard each $\sigma_i$
    as a group action
    $\sigma : \mathbb{Z_n} \to S_n = \mathrm{Perm}(S)$ where
    $\sigma(i)$ is the transposition of the $n$-th element with the
    $i$-th.
    $$
    |S| = \sum_{i=1}^n |\mathbb{Z}_n . i|
        = \sum_{i=1}^n [\mathbb{Z}_n : (\mathbb{Z}_n)_i]
    $$
    where
    $$
    (\mathbb{Z}_n)_i = \{ m \in \mathbb{Z}_n \mid \sigma(m)(i) = i \}
                    =
    $$
  }
\end{enumerate}

\end{proof}

\subsubsection{Cyclic Groups}
\begin{defn}[Cyclic Group]
$G$ is \emph{cyclic} when $\exists a \in G$ s.t.
$G = \langle a \rangle$.
\end{defn}

\begin{prop}
Cyclic groups are isomorphic to groups of integers.
\begin{enumerate}
  \item{If $G$ is infinite cyclic, then $G \simeq \mathbb{Z}$.}
  \item{If $G$ is finite cyclic, then $G \simeq \mathbb{Z}_{|G|}$.}
\end{enumerate}
\end{prop}
\begin{proof}
Pick a generator $a$ of $G$, so that
$G = \{ a^n | n \in \mathbb{Z} \}$, and construct a group homomorphism
$f : \mathbb{Z} \to G$ given by $k \mapsto a^k$. If $G$ is
infinite then $f$ is an isomorphism -- it is surjective since
$G = \langle a \rangle$ and injective since $G$ is infinite, so $a^k =
1$ iff. $k = 0$ and therefore $\ker f$ is trivial.

If $|G| = n < \infty$ then
there is instead an isomorphism $\bar{f} : \mathbb{Z}_n \to G$ given by
$k \mod n \mapsto a^k$. This is a group homomorphism since
$$
\bar{f}(k + l \mod n) = a^{k}a^{l} = f(k)f(l),
$$
surjective since $G = \{ 1, a, \dots, a^{k-1}\}$, and injective since
$$a^k = 1 \iff \exists m . k = mk \equiv 0 \mod n.$$
\end{proof}

\begin{prop}[Corollary of Lagrange's Theorem]
If $|G| = p$ for some $p$ prime, then $G$ is cyclic and so
$G \simeq \mathbb{Z}_p$.
\end{prop}
\begin{proof}
$|G| > 1$ so $\exists a \in G$ s.t. $a \neq 1$. Consider
$H = \langle a \rangle$. Then since $H < G$,
$|H|$ divides $|G| = p$ and
$|H| \neq 1$, so $|H| = |G|$ and thus $H = G$, so
$G = \langle a \rangle$ and $G$ is cyclic. Thus $G \simeq \mathbb{Z}_p$.
\end{proof}

This fact allows us to classify some groups (up to isomorphism) from
their order alone -- this is a major goal of group theory.

\subsection{Normal Subgroups}

\begin{defn}
Let $H < G$. Then $H$ is called \emph{normal} in $G$
($H \triangleleft G$) when $\forall a \in G$, $aH = Ha$.
\end{defn}

Equivalently,
\begin{itemize}
  \item{$\forall a \in G, a H a^{-1} = H$}
  \item{$\forall a \in G, a H a^{-1} \subset H$}
  \item{$\forall a \in G, a H \subset Ha$}
  \item{$\forall a \in G, a H a \subset a H$}
\end{itemize}
\begin{proof}

Starting from the original definition:
\begin{itemize}
  \item{
    Suppose $\forall a \in G$, $a H = H a$. Let $g \in a H a^{-1}$. Then
    $g = a h a^{-1} = h^\prime a a^{-1} = h^\prime \in H$, so $a H a^{-1} \subset H$.
    Let $h \in H$. Then $h = h a a^{-1} = a h^\prime a^{-1}$, so $h \in a H a^{-1}$.
  }
  \item{
    The forward implication from the first bullet to the second is clear.

    Suppose $\forall a \in G, a H a^{-1} \subset H$, and let $h \in H$. Then
    $a^{-1} h a \in a^{-1} H a \subset H$, so $a^{-1} h a \in H$, and then
    $h = a (a^{-1} h a) a^{-1} \in a H a^{-1}$, so $H \subset a H a^{-1}$.
  }
  \item{
    Suppose $\forall a$, $a H a^{-1} \subset H$, and let $ah \in aH$. Then
    $a h a^{-1} \in a H a \subset H$, so $a h a^{-1} \in H$. Then
    $ah = (aha^{-1})a \in Ha$, so $aH \subset Ha$.
  }
%  \item{Suppose $\forall a$, $aH \subset Ha$. First take $g \in Ha$ so that
%        $g = ha$ and note that
%        $$
%        a^{-1} g = a^{-1} h a = h^\prime a^{-1} a = h^\prime
%        $$
%        so that $g = a h^\prime$, and therefore $g \in aH$. Therefore we have
%        $aH = Ha$.
%
%        Next let $g \in aHa$ so $g = aha$. Then $g = a a h^
\end{itemize}
\end{proof}
%ote that, for instance
%begin{align*}
% H & = \{ x \in G | \exists h \in H . x = a h \}
%end{align*}
%o that
%begin{align*}
%^{-1} a H & = \{ x^\prime \in G | \exists x \in a H . x^\prime =
%^{-1} x = x a^{-1} a h = h \} \\
%         & = H,
%end{align*}
%o we may manipulate cosets algebraically.
%
%begin{align*}
%          a^{-1} H (a^{-1})^{-1} \subset H
% \implies a^{-1} H a \subset H  \\
% \implies H \subset a H a^{-1}, \\
%          a^{-1} H a \subset H
% \implies a^{-1} h a \in H \\
% \implies a (a^{-1} h a) a^{-1} \in a H a^{-1} \\
% \implies \forall h \in H . h \in a H a^{-1} \\
% \implies H \subset a H a^{-1}.
%end{align*}
%end{proof}

\begin{remark}
\begin{enumerate}
  \item{Normality is a relative notion to the ambient group.}
  \item{If $G$ is abelian, all subgroups are normal.}
\end{enumerate}
\end{remark}

\begin{prop}
Let $f : G \to G^\prime$ be a group homomorphism. Then
$\ker(f) \triangleleft G$.
\end{prop}
\begin{proof}
Show $a \ker(f) a^{-1} \subset \ker(f), \forall G$.
Let $x \in \ker(f).$ Then
$f(a x a^{-1}) = f(a) f(x) f(a)^{-1} = 1$, so
$a x a^{-1} \in \ker(f)$.
\end{proof}

\begin{xmpl}
\begin{enumerate}
  \item{$G = \mathrm{\mathrm{GL}}_n(\mathbb{R})$, $\det : G \to \mathbb{R}^\ast$.
        $\ker(\det) = \mathrm{SL}_n(\mathbb{R})
                    \triangleleft \mathrm{\mathrm{GL}}_n(\mathbb{R})$.
       }
  \item{The \emph{center} of $G$, $Z(G)$ is defined to be
        $$
        Z(G) = \{ z \in G | z g = g z, \forall g \in G \}.
        $$
        $Z(G) \triangleleft G$, since $g Z(G) g^{-1} \subset Z(G)$
        since $g z g^{-1} = z g g^{-1} = z \in Z(G)$. If
        $Z(G) = \{1\}$, then $G$ is said to be \emph{centerless}.
        }
\end{enumerate}
\end{xmpl}

% 9/10/2014

\begin{defn}[Affine Map]
An \emph{affine map} is a sum of linear maps with a translation. In
the case $\mathbb{R} \to \mathbb{R}$, the functions $f(x) = ax + b$,
$a \neq 0$ is affine.
\end{defn}

\begin{xmpl}[Normal Subgroup]
Let $G$ be the set of all non-constant \emph{affine} maps
$\mathbb{R} \to \mathbb{R}$ with operation $\circ$. That is,
$$
G = \{ f : \mathbb{R} \to \mathbb{R} | f(x) = ax + b,
                                       a \neq 0,
                                       a, b \in \mathbb{R}
    \}.
$$
For simplicity, set $L_{a,b}(x) = ax + b$. There are two standard
subgroups of $G$.

$A \triangleq \{ f | f(x) = ax, a \neq 0 \} = \{L_{a, 0} | a \neq 0\}$
is a subgroup of $G$ which is isomorphic to the multiplicative group
$\mathbb{R}^\ast$.

$T \triangleq \{ f | f(x) = x + b \} = \{ L_{1,b} | b \in \mathbb{R}
\}$ is a subgroup of $G$.

These subgroups are normal in $G$. Consider the map
$f : G \to \mathbb{R}^\ast \simeq A$ given by
$$
L_{a,b} \mathrel{\mathop{\mapsto}^{f}} a
$$.
This is a group homomorphism because
\begin{align*}
(L_{a,b} \circ L_{a^\prime, b^\prime})(x)
 & = L_{a,b}(L_{a^\prime, b^\prime}(x)) \\
 & = L_{a,b}(a^\prime x + b^\prime) \\
 & = a(a^\prime x + b^\prime) + b \\
 & = aa^\prime x + ab^\prime + b \\
 & = L_{a a^\prime, a b^\prime + b}(x),
\end{align*}
so $L_{a,b} \circ L_{a^\prime, b^\prime} = L_{a a^\prime, a b^\prime +
  b}$, and thus
$$
L_{a,b} \circ L_{a^\prime, b^\prime} \mathrel{\mathop{\mapsto}^{f}} a a^\prime,
$$
so this is a group homomorphism. The kernel is then
$\{L_{1,b} | b \in \mathbb{R} \} = T$, so $T$ is normal in $G$.
\end{xmpl}

Note that
$$
G \simeq \left\{
\left.\left(\begin{array}{c c}
a & b \\ 0 & 1
\end{array}\right)\right| a, b \in R, a \neq 0
\right\} < GL_2(\mathbb{R})
$$
by the group homomorphism
$$
L_{a,b} \mapsto \left(\begin{array}{c c}
a & b \\ 0 & 1
\end{array}\right).
$$

\subsection{Quotient Groups and Isomorphism Theorems}

\begin{defn}[Quotient Group]
Let $G$ be a group and $H \triangleleft G$, so that $aHa^{-1} = H$,
$\forall a \in G$, or equivalently $aH = Ha$.
The \emph{quotient group} is the set of all (left = right) cosets
$$
G / H = \{ aH | a \in G \}
$$
with a natural operation induced by the operation in the group:
$$
aH \cdot bH \triangleq aHbH
$$
or for $S, T \subset G$,
$$
S \cdot T \{ s \cdot t | s \in S, t \in T \}.
$$
This is indeed another coset because, by normality of $H$,
$aHbH = abHH = abH$. We need to check that it is associative, contains
identity ($H$), and contains all inverses ($(aH)^{-1} = a^{-1}H$).
Similarly there is the
\emph{canonical homomorphism} $\pi : G \to G/H$ given by
$\pi(a) = aH$. $\pi$ is surjective and
$\ker(\pi) = \{ a \in G | aH = H \} = H$, so $\pi$ is an isomorphism.
\end{defn}

Note that this quotient can also be written as a quotient by an
equivalence relation $G/\sim$ where
$$
a \sim b \triangleq aH = bH
$$
or by a function as $G/\pi$.

\subsection{Isomorphism Theorems}

\begin{theorem}[First Isomorphism Theorem]
Let $f : G \to G^\prime$ be a group homomorphism. There are two
associated subgroups $\mathrm{Im}(f) < G^\prime$ and
$\ker(f) \triangleleft G$. Then
$G / \ker(f) \simeq \mathrm{Im}(f)$, with an isomorphism given by
$a \ker(f) \mapsto f(a)$.
\end{theorem}
\begin{proof}
Set $\bar{f}(a \ker (f)) = f(a)$. We check that $\bar{f}$ is a
well-defined map $G / \ker(f) \to \mathrm{Im}(f)$ by showing that
$a \ker f = b \ker f \implies f(a) = f(b)$. Take
$a \ker f = b \ker f$ so that for all $k \in \ker f$ we have
$a k = b k^\prime$ for some $k^\prime$. But since $f$ is a homomorphism
we must have $f(1) = 1$ so that $1 \in \ker f$, and thus we have
$b^{-1} a = k^\prime \in \ker f$. Then
$$
b^{-1} a \in \ker(f)
  \implies f(b^{-1} a) = 1
  \implies f(b)^{-1} f(a) = 1
  \implies f(a) = f(b).
$$

Next, we prove $\bar{f}$ is a group homomorphism.
$$
\bar{f}(a \ker f \cdot b \ker f)
  = \bar{f} (a b \ker f)
  = f(ab)
  = f(a)f(b)
  = \bar{f}(a \ker f) \bar{f}(b \ker f)
$$
since $f$ is a homomorphism.

$\bar{f}$ is surjective by definition: let $y \in \mathrm{Im} f$. Then $y = f(x)$
for some $x \in G$, so $y = \bar{f}(x \ker f)$. Furthermore

\begin{align*}
\ker(\bar{f}) & = \{ a \ker f \mid \bar{f}(a \ker f) = f(a) = 1 \} \\
              & = \{ a \ker f \mid a \in \ker f \} \\
              & = \{ \ker f \},
\end{align*}
This means by definition that $\ker(\bar{f})$ has the same cardinality
as a singleton set so it must be trivial, thus $\bar{f}$ is injective.
\end{proof}

\begin{theorem}[Second Isomorphism Theorem]
Let $G$ be a group, $K \subset H$, and both $K$ and $H$ normal in
$G$. Then $K$ is normal in $H$ and we can consider $H / K$, which is
then a normal subgroup in $G / K$. Then $G / H \simeq (G / K) / (H / K)$.
\end{theorem}
\begin{proof}
$K \triangleleft G \iff \forall g \in G, g K g^{-1} = K$. Since
$H \subset G$, $h K h^{-1} = K$, so $K \triangleleft H$. Consider the
canonical maps $\pi_K : G \to G / K$ and
$\pi_H : G \to G / H$. We claim that there exists a homomorphism
$\varphi : G / K \to G / H$ given by $\varphi(gK) = gH$, and that
$\varphi \circ \pi_K = \pi_H$.

We show first that $\varphi \circ \pi_K = \pi_H$. We have
$$
(\varphi \circ \pi_K)(g) = \varphi(gK) = gH = \pi_H(g),
$$
and thus $\varphi = \pi_H \circ \pi_K^{-1}$, so $\varphi$ is a
composition of isomorphisms and therefore an isomorphism itself.

We observe next that
$$
\ker(\varphi)
  = \{ gK | gH = H \}
  = \{ gK | g \in H \}
  = H / K.
$$
Therefore
$\ker \varphi = H / K$, so by the first isomorphism theorem
$$
       (G / K) / (H / K)
=      (G / K) / \ker \varphi
\simeq \mathrm{Im} \varphi
=      G / H.
$$
\end{proof}

\begin{theorem}[Third Isomorphism Theorem]
Let $H < G$, $K < G$ such that $H$ \emph{normalizes} $K$, i.e.
$\forall h \in H, h K h^{-1} = K$ or
$\forall h \in H, h K = K h$. Then
$$
H \cap K \triangleleft H,
\quad HK < G,
\quad H / (H \cap K) \simeq HK / K.
$$
\end{theorem}
\begin{proof} There are three assertions to prove.
\begin{itemize}
  \item[($H \cap K \triangleleft H$.)]{
        $h \in H$, $\kappa \in H \cap K$. Then
        $h \kappa h^{-1} \in H$ since $h, \kappa \in H$
        and $h \kappa h^{-1} \in K$ since $h \in H$, $H$ normalizes
        $K$.
        Then $h \kappa h^{-1} \in H \cap K$, so $\forall h$
        $h (H \cap K) h^{-1} \subset H \cap K$ or equivalently
        $H \cap K \triangleleft H$.
       }
  \item[($HK < G$)]{
        $h_1 \kappa_1 h_2 \kappa_2
       = h_1 h_2 \kappa_1^\prime \kappa_2
       \in HK$, since $H$ normalizes $K$ and thus $Kh = hK$.
       This proves $HK$ is closed.

       $(h \kappa)^{-1} = \kappa^{-1} h^{-1} = h^{-1} \kappa^\prime \in HK$, since
       $H$ normalizes $K$ and thus $K h^{-1} = h^{-1} K$.

       Also $K \triangleleft HK$. Indeed,
       $$
       (h \kappa) \kappa^\prime (h \kappa)^{-1}
     = h (\kappa \kappa^\prime \kappa^{-1}) h^{-1} \in K,
       $$
       since $\kappa \kappa^\prime \kappa^{-1} \in K$ and
       $h K h^{-1} \subset K$.
       }
       \item[($H / H \cap K \simeq HK / K$)]{
             Consider $H \hookrightarrow HK \to^{\pi} HK / K$ that behaves as
             $$
             h \mapsto h \cdot 1 = h \mapsto h K.
             $$
             Notice that $\varphi = \pi \circ j$ is a homomorphism, and is
             surjective since $h \kappa K = h K$, and has the kernel
             $$
             \ker \varphi = \{ h \in H \mid hK = K \}
                          = \{ h \in H \mid h \in K \}
                          = H \cap K.
             $$
             Then $H / \ker \varphi \simeq \mathrm{Im} \varphi$, which
             is exactly
             $H / H \cap K \simeq HK / K$.
            }
\end{itemize}
\end{proof}

\begin{defn}[Normalizer]
Let $G$ be a group, $K < G$. Then the \emph{normalizer}
$N_K$ of $K$ in G is
$$
N_K = \{ g \in G \mid g K g^{-1} = K \}.
$$
\end{defn}

\begin{Problem}
Show that
\begin{itemize}
  \item{$K < N_K < G$.}
  \item{$K \triangleleft N_K$, and in fact $N_K$ is the largest
      subgroup of $G$ that contains $K$ such that $K$ is normal in
      that subgroup.}
\end{itemize}
\end{Problem}

\begin{Answer}
Certainly for $k, k^\prime \in K$ we have
$$
k K k^{-1} = (k k^\prime)K(k k^\prime)^{-1} = k^{-1} K k = K
$$
so $K < N_K$. Let $n_1, n_2 \in N_K$. Then
$$
n_1 n_2 k = n_1 k^\prime n_2 = k^{\prime\prime} n_1 n_2
$$
so $n_1 n_2 \in N_K$. Let $n \in N_K$. Then
$$
n^{-1} k = (k^{-1} n)^{-1} = (n k^\prime)^{-1} = (k^\prime)^{-1} n^{-1}
$$
so $n^{-1} \in N_K$. Since additionally $1$ commutes with all elements
and therefore is in the normalizer, we conclude that $K < N_K < G$.

$K \triangleleft N_K$ since for any
$n k n^{-1} \in n K n^{-1}$ we have
$$
n k n^{-1} = n n^{-1} k^\prime = k^\prime \in K
$$
so $n K n^{-1} \subset K$.

Next let $H < G$ be any subgroup such that $K \triangleleft H$.
Let $h \in H$. Then $h K = K h$, so $h \in N_K$. Then
$H \subset N_K$.
\end{Answer}

\begin{remark}
The assumption in the 3rd isomorphism theorem can be written $H < N_K$.
\end{remark}

\subsubsection{Special Case}
\begin{defn}[Semidirect Product]
Let $H < G$, $K \triangleleft G$, $H \cap K = \{ 1 \}$, and $HK = G$.
Then $G$ is the \emph{semidirect product} of $H$ and $K$ and we write
$G = H \ltimes K$ with the triangle pointing at $K$ to remember that
$K \triangleleft G$ (not that $K \triangleleft H$, since indeed we do
not have $K < H$ in our assumptions).

Since $K \triangleleft G$ we have $N_K = G$, so $H < N_K$ and
the third isomorphism theorem applies to give
$$
HK / K \simeq H / H \cap K \implies
G / K  \simeq H / \{ 1 \} \simeq H.
$$
\end{defn}

If $H$ normalizes $K$, then
$h \in H, k \in K$ gives $h k h^{-1} = k^\prime \in K$. This defnes a
group homomorphism $\psi : H \to \mathrm{Aut}(K)$ given by
$\psi(h)(k) = h k h^{-1} \in K$. We call $\psi(h)$ the
\emph{conjugation} by $h$. Now describe the group operation on
$G = HK$, using the operations of $H$ and $K$ and the homomorphism $\psi$.

Take $g_1, g_2 \in G$. Then $g_1 = h_1 k_1$, $g_2 = h_2 k_2$ Is this
unique? Can we write $g = hk = h^\prime k^\prime$? This would mean
$$
hk = h^\prime k^\prime
  \implies k = h^{-1} h^\prime k^\prime
  \implies k (k^\prime)^{-1} = h^{-1} h^\prime,
$$
but since $H \cap K$ is trivial this means
$hk = h^\prime k^\prime$.

$$g_1 \cdot g_2 = h_1 k_1 h_2 k_2 = h_1 h_2 k_1^\prime k_2$$
because $k_1 h_2 = h_2 k_1^\prime$ because $kH = Hk$, so
\begin{align*}
h_2^{-1} k_1 h_2 = k_1^\prime
& \implies
  \psi(h_2^{-1})(k_1) = k_1^\prime \\
& \implies
  \psi(h_2)^{-1}(k_1) = k_1^\prime \\
& \implies g_1 g_2 = h_1 h_2 \psi(h_2)^{-1} (k_1) k_2.
\end{align*}

% 9/15

We can express this construction differently. Start with
$H, K$ groups as given before
and a homomorphism $\psi : H \to \mathrm{Aut}(K)$.
Consider the direct product
$$
K \times H = \{(k, h) | k \in K, h \in H \}
$$
as a set but with the operation
$$
(k_1, h_1) \cdot (k_2, h_2) = (k_1 \psi(h_1)(k_2), h_1 h_2).
$$
Call this $G$.

\begin{Problem}
Show that $G$ is a group with this operation. This is an alternative
definition of the semidirect product $K \rtimes_\psi H$. Show also
that the previously defined $G$ is isomorphic to this. (Exercise 12 in
Lang).
\end{Problem}

\begin{Answer}
This is covered by Problems \#1 and \#2 in Homework \#2
-- the isomorphism is given by $(k, h) \mapsto kh$
which is a homomorphism because of the assumption that $H < N_K$.
\end{Answer}

\begin{xmpl}[Semidirect Products]
\begin{enumerate}
  \item{$G = S_3$, the symmetric group in three letters (permutations
      of $\{1,2,3\}$. We have elements $\sigma = (1 2 3)$ and
      $\tau = (1 2)$ where these notations indicate $1 \to 2 \to 3 \to
      1$ and $1 \to 2 \to 1, 3 \to 3$, respectively. Consider
      $K = \langle \sigma \rangle = \{1, \sigma, \sigma^2\}$ and
      $H = \langle \tau \rangle = \{1, \tau\}$. Then
      $$
      G = \{1, \sigma, \sigma^2, \tau, \tau \sigma, \tau \sigma^2\}
        = H \ltimes K.
      $$
      The conjugation
      $\psi : \langle \tau \rangle \to Aut(\langle \sigma \rangle)$
      is given by
      $$
      \psi(\tau)(\sigma) = \tau \sigma \tau^{-1}
                         = \sigma^2 \tau \tau^{-1}
                         = \sigma^2,
      $$
      because $\tau \sigma = \sigma^2 \tau$ because
      $$
      (1 2) \cdot (1 2 3) = (2 3), \quad
      (1 2 3) \cdot (1 2 3) \cdot (1 2)
        = (1 3 2) \cdot (1 2) = (2 3).
      $$
      We can also realize that these subgroups are cyclic
      groups of order 2 and 3, or
      $$
      \langle \tau \rangle \simeq \mathbb{Z}_2, \quad
      \langle \sigma \rangle \simeq \mathbb{Z}_3
      $$
      and
      $$
      \psi : \mathbb{Z}_2
           \to \mathrm{Aut}(\mathbb{Z}_3)
           =   \mathbb{Z}_3^* \simeq \mathbb{Z}_2
      $$
      so $\psi$ is identity under these identifications.
      Then we can rewrite
      $$
      S_3 \simeq \mathbb{Z}_3 \rtimes_\psi \mathbb{Z}_2.
      $$
      }
  \item{Let $G = \mathrm{GL}_2(\mathbb{R})$. Consider
        $K = \mathrm{SL}_2(\mathrm{R})$ and
        $\mathrm{R}^* = \mathrm{R} - \{ 0 \}.$
        Recall that
        $\mathrm{SL}_2(\mathbb{R}) \triangleleft
        \mathrm{GL}_2(\mathbb{R})$
        because $\mathrm{SL}_2 = \ker \det$.
        We can embed
        $\mathbb{R}^* \hookrightarrow \mathrm{GL}_2(\mathbb{R})$
        by $a \mapsto \left(\begin{array}{c c} a & 0 \\ 0 &
            1\end{array}\right)$ and call this subgroup $H$.

        Note that $H \cap K = \{ 1 \}$ and
        $\mathrm{GL}_2(\mathbb{R}) = H K$ because
        given $A \in \mathrm{GL}_2(\mathbb{R})$ we can write
        $$
        A = \left(\begin{array}{c c}
              \det A & 0 \\ 0 & 1
            \end{array}\right) \cdot A^\prime
        $$
        with $\det A^\prime = 1$.

        $$
        \psi : \mathbb{R}^*
           \to \mathrm{Aut}(\mathrm{SL}_2(\mathbb{R})
        $$
        is given by
        $$
        \psi(a)(A) =
          \left(\begin{array}{c c}
            a & 0 \\ 0 & 1
          \end{array}\right) \cdot A \cdot
          \left(\begin{array}{c c}
            a & 0 \\ 0 & 1
          \end{array}\right)^{-1}.
          $$
       }
\end{enumerate}
\end{xmpl}

\subsection{Decomposition of Groups}

\begin{defn}[Tower of Subgroups]
Let $G$ be a group. A sequence of subgroups
$$
G = G_0 > G_1 > \cdots > G_i > \cdots
$$
will be called a \emph{tower} of subgroups of $G$.
It is called a \emph{normal tower} if
$\forall i, G_{i+1} \triangleleft G_i$, an
\emph{abelian tower} if it is a normal tower such that
$G_{i} / G_{i+1}$ is abelian, and a \emph{cyclic tower}
when $G_{i} / G_{i+1}$ is cyclic.
\end{defn}

\begin{defn}[Solvable Group]
A group is called \emph{solvable} if it has a finite
abelian tower ending in $\{ 1 \}$.
\end{defn}

This terminology comes from a Galois group being solvable, in which
case corresponding polynomials can be solved by radicals.

\begin{theorem}[Feit-Thompson]
  Every finite group of odd order is solvable.
\end{theorem}

\begin{xmpl}
\begin{enumerate}
  \item{Consider the subgroup $G < \mathrm{GL}_3(\mathbb{R})$ given by
      $$
      G = \left\{\left.
          \left(\begin{array}{c c c}
            a & d & e \\ 0 & b & f \\ 0 & 0 & c
          \end{array}\right)\right|a, b, c \neq 0\right\},
      $$
      the \emph{parabolic subgroup}. $G$ is solvable.

      \begin{proof}
        Consider $\phi : G
                     \to \mathbb{R}^* \times \mathbb{R}^* \times \mathbb{R}^*$
        given by
        $$
        \left(\begin{array}{c c c}
          a & d & e \\ 0 & b & f \\ 0 & 0 & c
        \end{array}\right) \mapsto (a,b,c).
        $$
        This is a homomorphism (exercise) which is
        surjective. Take $\ker \varphi = G_1 \triangleleft G$.
        This $\ker \varphi$ is the matrices which have ones on the
        diagonal. Then the quotient, by the first isomorphism theorem, is
        $$
        G / \ker \varphi
          \simeq \mathrm{Im}(\varphi)
          = \mathbb{R}^* \times \mathbb{R}^* \times \mathbb{R}^*.
        $$
        Note that this image is abelian.

        Next take
        $$
         G_2 =
          \left\{\left.
          \left(\begin{array}{c c c}
            1 & 0 & x \\ 0 & 1 & 0 \\ 0 & 0 & 1
          \end{array}\right)\right| x \in \mathbb{R}\right\}.
        $$
        Check that $G_2$ is normal in $G_1$, and that
        $G_1 / G_2 \simeq \mathbb{R} \times \mathbb{R}$, since it is
        determined by the two zeroed entries.

        Observe that $G_2 \simeq \mathbb{R}$ by the map taking its
        top-right element. So we have the abelian tower
        $$
        \left(\begin{array}{c c c}
          a & d & f \\ 0 & b & e \\ 0 & 0 & c
        \end{array}\right)
        \triangleright
        \left(\begin{array}{c c c}
          1 & d & f \\ 0 & 1 & e \\ 0 & 0 & 1
        \end{array}\right)
        \triangleright
        \left(\begin{array}{c c c}
          1 & 0 & x \\ 0 & 1 & 0 \\ 0 & 0 & 1
        \end{array}\right)
        \triangleright \{ 1 \}.
        $$
        This is true for any field.
      \end{proof}
    }
\end{enumerate}
\end{xmpl}

% 9/17

\begin{prop}
Let $G$ be a group and $H \triangleleft G$. Then $G$ is solvable if
and only if both $H$ and the quotient $G / H$ are solvable.
\end{prop}
\begin{proof}

We show the reverse implication first.
\begin{itemize}
\item[($\impliedby$)]
{
Assume $H, G / H$ are solvable. Then there are towers
$$
H = H_0 \triangleright \cdots \triangleright H_n = \{ 1 \}, \quad
H_i/H_{i+1} \text{ abelian}
$$
$$
G/H = K_0 \triangleright \cdots \triangleright K_m = \{ 1 \}, \quad
K_i/K_{i+1} \text{ abelian}.
$$
Confirm that the inverse image $f^{-1}(K_i)$ of a normal subgroup is
normal under any group homomorphism $f$. Let $x \in f^{-1}(N)$ for
some homomorphism $f : G \to G^\prime$ and some $N \triangleleft
G^\prime$. We have that
$$
  f(g x g^{-1})
= f(g) f(x) f(g)^{-1}
= f(g) f(g)^{-1} n = n \in N
$$
and so $g x g^{-1} \in f^{-1}(N)$. Therefore
$g f^{-1}(N) g = f^{-1}(N)$.

Note that
$$
\pi^{-1}(K_i) / \pi^{-1}(K_{i+1}) \simeq K_i / K_{i+1}
$$
since by restricting $\pi$ we have a surjective map
$\pi|_{\pi^{-1}(K_i)} : \pi^{-1}(K_i) \to K_i$ with kernel $H$,
so
$$
\pi^{-1}(K_i) / H \simeq K_i.
$$
Now since $H \triangleleft G$ and as just shown
$\pi^{-1}(K_i) \triangleleft G$, we apply the second isomorphism
theorem to see
$$
K_i / K_{i+1} \simeq (\pi^{-1}(K_i) / H) / (\pi^{-1}(K_{i+1}) / H)
             \simeq \pi^{-1}(K_i) / \pi^{-1}(K_{i+1}).
$$
So $\forall i$, $\pi^{-1}(K_i) / \pi^{-1}(K_{i+1})$ are all abelian,
and since $\pi^{-1}(K_i) \triangleleft G$ we have therefore formed
an abelian tower.
}
\item[($\implies$)]
{
Next, assume $G$ is solvable and show $H, G/H$ are solvable.
We may form the tower
$$
H \triangleright H \cap G_1
  \triangleright H \cap G_2
  \triangleright \cdots
  \triangleright H \cap G_n = \{ 1 \}
$$
since
\begin{enumerate}
  \item{
    $$
    \forall i . H \cap G_{i+1} \triangleleft H \cap G_i.
    $$
    Take $h \in H \cap G_i$ and $g \in H \cap G_{i+1}$.
    Then since $g \in G_{i+1} \triangleleft G_i$ and
    $h \in G_i$ we have $h g h^{-1} \in G_{i+1}$ and
    $h, g \in H$, so $h g h^{-1} \in H \cap G_{i+1}$.
  }
  \item{
    Construct a homomorphism
    $$
    \varphi : (H \cap G_i) / (H \cap G_{i+1}) \to G_i / G_{i+1}
    $$
    given by
    $$
    \varphi(h \mod H \cap G_{i+1}) \triangleq h \mod G_{i+1},
    $$
    which is injective since if $h \in H \cap G_i$,
    $\varphi(h \mod H \cap G_{i+1}) = 1 \mod G_{i+1}$ then
    $h \in H \cap G_{i+1}$ so
    $h \mod H \cap G_{i+1} = 1,$
    thus $\ker \varphi$ is trivial.
    Check that this is well-defined (show that it is independent of
    the representative).

    $\varphi$ gives an isomorphism between
    $H \cap G_i / H \cap G_{i+1}$ and a subgroup of the abelian group
    $G_i / G_{i+1}$, so $H \cap G_i / H \cap G_{i+1}$ is abelian.
  }
\end{enumerate}
Similarly we can show that $G / H$ is solvable (exercise).
}
\end{itemize}
\end{proof}

Let $H \triangleleft G$. Consider the quotient $G / H$. Then we have
two groups ``smaller'' than the original group. We can then decompose
our understanding of $G$ by considering $H$ and $G/H$ similarly.

\begin{defn}[Simple Group]
A group that has no normal proper subgroup is called \emph{simple},
i.e. if the only normal subgroups of $G$ are $\{1\}$ and $G$.
\end{defn}

\begin{theorem}[Jordan-H\"older Theorem]
Let $G$ be a finite group. Suppose
$$
G = G_0 \triangleright G_1
        \triangleright \cdots
        \triangleright G_r = \{1\}
$$
and
$$
G = H_0 \triangleright H_1
        \triangleright \cdots
        \triangleright H_s = \{1\}
$$
are two normal towers such that
$G_i / G_{i+1}$ and $H_j / H_{j+1}$ are simple groups
$\forall i, j$. Then $r = s$ and there is a permutation
on the indices $i \to i^\prime$ such that
$G_i / G_{i+1} \simeq H_{i^\prime} / H_{i^\prime + 1}, \forall
i$. This is analogous to prime factorization of positive integers.
\end{theorem}

\begin{remark}
Every finite group admits a normal tower ending at $\{ 1 \}$ such that
the quotients are simple, because a group $G$ is either simple itself
or has a nontrivial normal subgroup by definition, and thus admits a
refinement. This procedure will stop eventually because the group is finite.
\end{remark}

\begin{defn}[Refinement]
A tower of $G$ is called a \emph{refinement} of another tower of $G$
when it is obtained by inserting additional subgroups.
\end{defn}

\begin{defn}[Equivalent Towers]
Two normal towers of $G$
$$
G = G_0 \triangleright G_1
        \triangleright \cdots
        \triangleright G_r = \{1\}
$$
and
$$
G = H_0 \triangleright H_1
        \triangleright \cdots
        \triangleright H_s = \{1\}
$$
are \emph{equivalent} if their quotients are isomorphic up to a permutation.
\end{defn}

\begin{theorem}[Schreier]
Any two normal towers of $G$ that both end with $\{1\}$ have
equivalent (isomorphic up to permutation) refinements.
\end{theorem}

\begin{proof}
Let
\begin{align*}
G &= G_0 \triangleright G_1 \triangleright
     \cdots \triangleright G_r = \{ 1\}, \\
G &= H_0 \triangleright H_1 \triangleright
     \cdots \triangleright H_s = \{ 1\}. \\
\end{align*}
Consider that
$$
H_0 \cap G_i > H_1 \cap G_i > \cdots > H_s \cap G_i = \{1\}.
$$
Define $G_{i,j} = G_{i+1}(H_j \cap G_i)$,
$H_{j,i} = (H_j \cap G_i)H_{j+1}$. Then
$$
G_{i+1} \subset G_{i+1} (H_j \cap G_i) \subset G_i
$$
for all $i=0,\dots,r$ and $j=0,\dots,s$.

It can be shown that these
constructions are both normal towers of $G$ that end in $\{1\}$ and
have the same number of steps (because we have the same number of
pairs for each tower). Furthermore $\{G_{i,j}\}$ is a refinement of
$\{G_i\}$ and $\{H_{j,i}\}$ is a refinement of $\{H_j\}$.

We claim that there are isomorphisms
$G_{i,j} / G_{i,j+1} \simeq H_{j,i} / H_{j, i+1}$. This implies the
theorem. Let's write it in all its disgusting glory. This says
$$
\frac{G_{i+1} (H_j \cap G_i)}{G_{i+1} (H_{j+1} \cap G_i)}
\simeq
\frac{(H_j \cap G_i) H_{j+i}}{(H_j \cap G_{i+1}) H_{j+1}}
$$
We then want to show that
$$
\frac{u(U \cap V)}{u(U \cap v)}
\simeq
\frac{U \cap V}{(u \cap V)(U \cap v)}
\simeq
\frac{(V \cap U)v}{(V \cap u)v},
$$
(see diagram) where

$$
U = G_i, \quad
u = G_{i+1}
V = H_j, \quad
v = H_{j+1}.
$$

We proceed by the third isomorphism theorem
$$
HK / K \simeq H / (H \cap K)
$$
if $H$ normalizes $K$. We set first
$H = U \cap V$ and $K = u(U \cap v)$, then
$K = (V \cap u)v$. Check that
\begin{itemize}
  \item{$H$ normalizes $K$,}
  \item{$HK$ is the numerator we want,}
  \item{and $H \cap K = (u \cap V)(U \cap v)$.}
\end{itemize}

This is also called the \emph{Butterfly lemma}.
\end{proof}

This immediately implies the Jordan-H\"older theorem because a normal
tower with simple quotients has no nontrivial refinements.

\end{document}
& \implies g_1 g_2 = h_1 h_2 \psi(h_1)^{-1} (k_1) k_2.
\end{align*}

\end{document}
